---
title: "Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Framework
```{r file and summary}

data <- read.csv("Kickstarter.csv",na.strings=c("","NA"))

sum<-summary(data)

```

1) Set your hypothesis, have a specific goal.
# Success of project based on backers, money raised, and category. (Quanitative variables are)
# For hyp testing (how do these attributes effect the data)

# Project with #(of backers) is successful, or 
# Average project in technology raises #($) or
# Average successful project has (# of backers) or
# ....... (this can be whatever u guys want.)

```{r hypothesis}



```

# 2)	Review your data to ensure that they are appropriate and complete and can help you prove or disprove your hypothesis.

# we want to only base our research on project on USD income.

```{r data review}

data<-data[data$currency == "USD",]

```

# 3)	Complete literature review on the subject/hypothesis and determine if there is any relevant research/study has been already completed. Study the literature and include citations in your final report.

https://towardsdatascience.com/predicting-the-success-of-kickstarter-campaigns-3f4a976419b9


# 4)	Once you ensure that your data are sufficient and you can initially rely on them to run your modeling, clean them up.
```{r clean up}

dataR<-data[c(1,3,5,6,9,17,23,27,33,37)]

parseType<-function(test){
test<-gsub('"', "",test)
sub(".*:name",",",test)
test<-sub(".*,name:","",test)
sub(",.*","",test)
}

dataR$category<-lapply(data$category,parseType)
dataR$location<-lapply(data$location,parseType)

```

Exploratory Analysis
# 5)	Run Exploratory analysis on at least 5 total variables where 2 of them are quantitative.

# Varialbles we want to focus on / analyze :
# (1) backers_count , (2) ?category? , (3) goal [$]
# (4) pledged [amount of people supporting], (5) usd_pledged [total raised]

# 5a. Check for missing data
```{r missing data check}

dataFinal<-dataR[rowSums(is.na(dataR))==0,]

```

# 5b.	Check for outliers, IQR, and summarize the statistics.
```{r outliers, IQR, and statistic summaries}

summary(dataFinal$backers_count)
boxplot.stats(dataFinal$backers_count)$out
IQR(dataFinal$backers_count)

summary(dataFinal$goal)
boxplot.stats(dataFinal$goal)$out
IQR(dataFinal$goal)

summary(dataFinal$pledged)
boxplot.stats(dataFinal$pledged)$out
IQR(dataFinal$pledged)

summary(dataFinal$usd_pledged)
boxplot.stats(dataFinal$usd_pledged)$out
IQR(dataFinal$usd_pledged)

```

# 5c.	Disect your variables in a way that will help you with your analysis.
```{r variable dissection}

typeof(dataFinal$backers_count)

typeof(dataFinal$goal)

typeof(dataFinal$category)

typeof(dataFinal$pledged)

typeof(dataFinal$usd_pledged)

```

# 5d.	Determine the distribution ( if any that your data follow, experimentally and theoretically)
```{r distribution}

par(mfrow = c(1, 4)) #Creates a 1 x 4 picture of the graphs

plot(dataFinal$backers_count)

plot(dataFinal$goal)

plot(dataFinal$pledged)

plot(dataFinal$usd_pledged)

cat("The quantative variables do not follow any distributions.")

```

# 5e. Show your analysis in both tables/charts and visually (histograms, qqnorm plots, boxplots etc.Statistical Modeling)
```{r tables/charts and graphs}

par(mfrow = c(3,3)) #Creates a 3 x 3 picture of the graphs

#Backer Count
hist(dataFinal$backers_count, main = "Backers", xlab = "Backer Count")
boxplot(dataFinal$backers_count)
qqnorm(dataFinal$backers_count)
qqline(dataFinal$backers_count)

#Goal
hist(dataFinal$goal, main = "Goal")
boxplot(dataFinal$goal)
qqnorm(dataFinal$goal)
qqline(dataFinal$goal)

#Pledged
hist(dataFinal$pledged, main = "Pledged")
boxplot(dataFinal$pledged)
qqnorm(dataFinal$pledged)
qqline(dataFinal$pledged)

par(mfrow = c(1, 3)) #Creates a 2 x 3 picture of the graphs

#USD Pledged
hist(dataFinal$usd_pledged, main = "USD Pledged")
boxplot(dataFinal$usd_pledged)
qqnorm(dataFinal$usd_pledged)
qqline(dataFinal$usd_pledged)

```

# 6) Determine correlations, do correlation comparisons ( technically and visually ) use both plots or pairs for your graphical representations. Split your graphs in ways that will help you to conclude and infer based on your model.
```{r correlations}



```

# 7) Your model and main hypothesis should be answered either using ANOVA or Regression Analysis, or both. - This may mean, depending on your data, that you may need to use a categorical variable to dissect your data, and that you may need to have data with many more than just 2 quantitative variables.

```{r Anove/Regression Analysis}



```

#Report
The report should be written in R Markdown and then transformed to HTML.  Your report needs to have the following sections:
  
  Introduction - In this section, explain briefly the purpose of your analysis. Identify your hypothesis, and in a single sentence refer to the results of your work.
 	
  Data - a section describing the data set and how you loaded and transformed it in R.  Include R code blocks within your comments and explain what the code is doing.
 	
 	Analysis - walk through the analysis that you performed. Include R code blocks within your comments and explain what the code is doing.

	Issues - Refer to any issues you had with collecting your data, cleaning your data, or implementing your model.

	Results - any plots, tables, or other results which gave you the answers to the questions that you were asked. Include R code blocks for any plots and explain what the code is doing.

	Discussion - a brief (one or two paragraphs) discussion of the results and how they validate or not your initial claims. 

```{r space}

#Just here to add space

```

# Presentation
You should prepare a brief (5-8 minutes) presentation on your project

You should create a Powerpoint presentation on your report.

Your Presentation should be narrated, or you should create a recorded video, presenting the Powerpoint.

#The presentation should include:
   	Introduction - quick reference to the problem you are analyzing and your data
  	Hypothesis - main claim or claims you worked on
  	Methodology - what statistical tools you used to complete your analysis and why?
   	Issues - any issues you had while you run your analysis and how did you resolve them
  	Tables, Charts and Graphs and measures with explanation for your EDA
  	Results of Correlation, Regression, and ANOVA analysis
  	Conclusions
  	Future questions and further analysis not addressed in this project
  	References

















