---
title: "Final Project"

output:
  html_document: default
  pdf_document: default
---
<style type="text/css">
.main-container {max-width: 1800px;margin-left: auto;margin-right: auto;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "140%")
```

# Framework
```{r file and summary}

library(arsenal) # used for compare

#Read in big data set
data <- read.csv("KickstarterBig.csv",na.strings=c("","NA"))

```
# 1) Set your hypothesis, have a specific goal.
# Success of project based on backers, money raised, and category. (Quanitative variables are)
# For hyp testing (how do these attributes effect the data)

# Project with #(of backers) is successful, or 
# Average project in technology raises #($) or
# Average successful project has (# of backers) or


```{r hypothesis}

# ....... (this can be whatever u guys want.)


```

# 2)	Review your data to ensure that they are appropriate and complete and can help you prove or disprove your hypothesis.

# we want to only base our research on project on USD income and have finished.

```{r data review}
# We want to focus on US data only that have completed the project deadline.
data<-data[data$currency == "USD",]
data<-data[data$state != "live",]
data<-data[data$state != "canceled",]

#Lets assume projects under goal of 100$ USD are innacurate data of reasonable project.

data<-data[data$goal > 100,]

```

# 3)	Complete literature review on the subject/hypothesis and determine if there is any relevant research/study has been already completed. Study the literature and include citations in your final report.

# https://towardsdatascience.com/predicting-the-success-of-kickstarter-campaigns-3f4a976419b9


# 4)	Once you ensure that your data are sufficient and you can initially rely on them to run your modeling, clean them up.
```{r clean up}
dataR<-data[-1]
dataR<-dataR[c(1,3,5,9,17,23,27,33)]

parseCateg<-function(test){
test<-gsub('"', "",test)
sub(".*:parent_name",",",test)
test<-sub(".*slug:","",test)
sub(",.*","",test)
test<-gsub("/.*","",test)
test<-gsub(",.*","",test)
}

parseLoc<-function(test){
  test<-gsub('"', "",test)
  sub(".*:name",",",test)
  test<-sub(".*,name:","",test)
  sub(",.*","",test)
}


dataR$category<-lapply(data$category,parseCateg)
dataR$location<-lapply(data$location,parseLoc)


UniqueCat <- unique(dataR$category)
print("The categories:",str(UniqueCat))

# Factorize descriptions

dataR <- as.data.frame(lapply(dataR, unlist))
dataR$category <- as.factor(dataR$category)

```

Exploratory Analysis
# 5)	Run Exploratory analysis on at least 5 total variables where 2 of them are quantitative.

# Varialbles we want to focus on / analyze :

# (1) backers_count
# (2) category of project
# (3) goal amount [USD for project]
# (4) pledged [amount of people supporting]
# (5) usd_pledged [total raised]

# 5a. Check for missing data
```{r missing data check / more prelims }

# Clean data
dataFinal <- dataR[rowSums(is.na(dataR))==0,]
# make the success / fail samples and tables for prelim observations.

dataFinal$pReached<- ifelse(dataFinal$state == "successful",( (dataFinal$pledged / dataFinal$goal) ),
                                                            ( (dataFinal$goal - dataFinal$pledged) / dataFinal$goal) )
# convert to percentages
dataFinal$pReached<- ifelse(dataFinal$state == "successful", ( dataFinal$pReached * 100 ), ( dataFinal$pReached / 100 ))

# Summarise  the two different data sets to see which projects succeed the most or fail the most
Succ<-dataFinal[dataFinal$state == "successful",]
Fail<-dataFinal[dataFinal$state == "failed",]

table(dataFinal$state)

#summary(comparedf(Succ, Fail))

```

# 5b.	Check for outliers, IQR, and summarize the statistics.
# TODO : Make actual plots, and list the outliers ( how many, what are they, where are they from)

```{r outliers, IQR, and statistic summaries}

cat("--Summary of backers count--")
summary(dataFinal$backers_count)
cat("Outliers count: ",length(sort(boxplot.stats(dataFinal$backers_count)$out)))
cat("IQR: ", IQR(dataFinal$backers_count))

par(mfrow = c(1, 2))
boxplot(dataFinal$backers_count,xlab="Full backers boxplot")
boxplot(dataFinal$backers_count,xlab="Backers without outliers",outline=FALSE)

 
cat("--Summary of goal $--")
summary(dataFinal$goal)
cat("Outliers count: ",length(sort(boxplot.stats(dataFinal$goal)$out)))
cat("IQR: ", IQR(dataFinal$goal))

par(mfrow = c(1, 2))
boxplot(dataFinal$goal,xlab="Full goal $ boxplot")
boxplot(dataFinal$goal,xlab="Goal $ without outliers",outline=FALSE)


cat("--Summary of pledged $--")
summary(dataFinal$pledged)
cat("Outliers count: ",length(sort(boxplot.stats(dataFinal$pledged)$out)))
cat("IQR: ", IQR(dataFinal$pledged))

par(mfrow = c(1, 2))
boxplot(dataFinal$pledged,xlab="Full pledged $ boxplot")
boxplot(dataFinal$pledged,xlab="Full pledged $ without outliers",outline=FALSE)


cat("--Summary of % raised--")
summary(dataFinal$pReached)
cat("Outliers count: ",length(sort(boxplot.stats(dataFinal$pReached)$out)))
cat("IQR: ", IQR(dataFinal$pReached))

par(mfrow = c(1, 2))
boxplot(dataFinal$pReached,xlab="Full % raised boxplot")
boxplot(dataFinal$pReached,xlab="% raused without outliers",outline=FALSE)


```

# 5c.	Disect your variables in a way that will help you with your analysis.
```{r variable dissection}

# ----- lets categorise the data by type of project ----- #

barplot(table(dataFinal$category),col= rainbow(15),las = 2,cex.names = .85,horiz = FALSE,beside=TRUE, xpd = FALSE, main="Project category popularity", ylab="Number of projects")


```

# 5d.	Determine the distribution (if any that your data follow, experimentally and theoretically)
# TODO  : Make this pretty, with labels, colors, ect, on graphs, and then make summaries of data here so we can
# Transfer that directly onto our presentation when we are making the powerpoint

```{r distribution}

par(mfrow = c(1, 4)) #Creates a 1 x 4 picture of the graphs

plot(dataFinal$backers_count, ylab = "# of Backers", main = "Backer Count", col = "deepskyblue3")
plot(dataFinal$goal, ylab = "Goals in $", main = "Monetary Goal", col = "purple3")
plot(dataFinal$pledged, ylab = "Amount pledged", main = "Pledged", col = "indianred3")
plot(dataFinal$pReached, ylab = "Amount Reached to goal", main = "USD Pledged", col = "goldenrod1")

```

#The quantative variables do not follow any distributions.

```{r summary}

summary(dataFinal$backers_count)
summary(dataFinal$goal)
summary(dataFinal$pledged)
summary(dataFinal$dataFinal$pReached)

```

# 5e. Show your analysis in both tables/charts and visually (histograms, qqnorm plots, boxplots etc.Statistical Modeling)
# TODO : analysis 

```{r tables/charts and graphs}

par(mfrow = c(3,3)) #Creates a 3 x 3 picture of the graphs

#Backer Count
hist(dataFinal$backers_count, xlab = "# of Backers", main = "Backer Count", col = "deepskyblue3")
boxplot(dataFinal$backers_count, ylab = "# of backers", main = "Backer Count", col = "deepskyblue3")
qqnorm(dataFinal$backers_count, ylab = "# of backers", main = "Backer Count", col = "deepskyblue3")
qqline(dataFinal$backers_count)

#Goal
hist(dataFinal$goal, xlab = "Goal in $", main = "Monetary Goal", col = "purple3")
boxplot(dataFinal$goal, ylab = "Goal in $", main = "Monetary Goal", col = "purple3")
qqnorm(dataFinal$goal, ylab = "Goal in $", main = "Monetary Goal", col = "purple3")
qqline(dataFinal$goal)

#Pledged
hist(dataFinal$pledged, xlab = "Amount Pledged", main = "Pledged", col = "indianred3")
boxplot(dataFinal$pledged, ylab = "Amount Pledged", main = "Pledged", col = "indianred3")
qqnorm(dataFinal$pledged, ylab = "Amount Pledged", main = "Pledged", col = "indianred3")
qqline(dataFinal$pledged)

par(mfrow = c(1, 3)) #Creates a 1 x 3 picture of the graphs

#USD Pledged
hist(dataFinal$pReached, xlab = "Amount Pledged in USD", main = "USD Pledged", col = "goldenrod1")
boxplot(dataFinal$pReached, ylab = "Amount Pledged in USD", main = "USD Pledged", col = "goldenrod1")
qqnorm(dataFinal$pReached, ylab = "Amount Pledged in USD", main = "USD Pledged", col = "goldenrod1")
qqline(dataFinal$pReached)

```

# 6) Determine correlations, do correlation comparisons (technically and visually ) use both plots or pairs for your graphical representations. Split your graphs in ways that will help you to conclude and infer based on your model.
```{r adding an id to the items under category, include=FALSE}

unique(dataFinal$category)
dataFinal <- transform(dataFinal, id = match(category, unique(category))) 

```

The portion above adds a new column labeled "id". Every unique item under category was given a unique id.

```{r scatter plots}
# TODO: Change to by category or leave as ID.

par(mfrow = c(1, 4))#Creates a 1 x 4 picture of the graphs

plot(dataFinal$category, dataFinal$backers_count, xlab = "", ylab = "Backer Count", main = "Category vs Backer Count", col = "deepskyblue3",las =2, cex.names = .5)

plot(dataFinal$category, dataFinal$goal, xlab = "", ylab = "Goal", main = "Category vs Goal", col = "purple3",las =2,cex.names = .5)


plot(dataFinal$category, dataFinal$pledged, xlab = "", ylab = "Pledged", main = "Category vs Pledged", col = "indianred3",las =2,cex.names = .5)

plot(dataFinal$category, dataFinal$pReached, xlab = "", ylab = "Percentage Raised",cex.names = .25, main = "Category vs % Raised", col = "goldenrod1",las =2,beside=TRUE, xpd = FALSE)

```

```{r scatter plot matrix}

# fix it to be by category

pairs(~dataFinal$backers_count + dataFinal$goal + dataFinal$pledged + dataFinal$pReached +
       dataFinal$category, main = "Backer Count, Goal, Pledged, USD Pledged vs Category")

```

```{r correlation}

cor(as.numeric(dataFinal$category), dataFinal$backers_count)

cor(as.numeric(dataFinal$category), dataFinal$goal)

cor(as.numeric(dataFinal$category), dataFinal$pledged)

cor(as.numeric(dataFinal$category), dataFinal$pReached)

# The correlation of all four attributes are very close to 0. This shows that there is minimal to none relationship with the category.

```

# 7) Your model and main hypothesis should be answered either using ANOVA or Regression Analysis, or both. - This may mean, depending on your data, that you may need to use a categorical variable to dissect your data, and that you may need to have data with many more than just 2 quantitative variables.

```{r Anova/Regression Analysis}



```

#Report
The report should be written in R Markdown and then transformed to HTML.  Your report needs to have the following sections:
  
  Introduction - In this section, explain briefly the purpose of your analysis. Identify your hypothesis, and in a single sentence refer to the results of your work.
 	
  Data - a section describing the data set and how you loaded and transformed it in R.  Include R code blocks within your comments and explain what the code is doing.
 	
 	Analysis - walk through the analysis that you performed. Include R code blocks within your comments and explain what the code is doing.

	Issues - Refer to any issues you had with collecting your data, cleaning your data, or implementing your model.

	Results - any plots, tables, or other results which gave you the answers to the questions that you were asked. Include R code blocks for any plots and explain what the code is doing.

	Discussion - a brief (one or two paragraphs) discussion of the results and how they validate or not your initial claims. 

```{r space}

#Just here to add space

```

# Presentation
You should prepare a brief (5-8 minutes) presentation on your project

You should create a Powerpoint presentation on your report.

Your Presentation should be narrated, or you should create a recorded video, presenting the Powerpoint.

#The presentation should include:
   	Introduction - quick reference to the problem you are analyzing and your data
  	Hypothesis - main claim or claims you worked on
  	Methodology - what statistical tools you used to complete your analysis and why?
   	Issues - any issues you had while you run your analysis and how did you resolve them
  	Tables, Charts and Graphs and measures with explanation for your EDA
  	Results of Correlation, Regression, and ANOVA analysis
  	Conclusions
  	Future questions and further analysis not addressed in this project
  	References


















